Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-22 ICML-submission/appendix-nearest-neighbor-algorithm.tex
Fixes

\section{Nearest neighbor algorithm}
\begin{algorithm}
\caption{A nearest-neighbor algorithm for \textsc{Online Multiclass Linear Classification with
Bandit Feedback}}
\begin{algorithmic}[1]
\STATE $S \gets \emptyset$.
\FOR{$t=1,2,\ldots,T$:}
\IF{$\min_{(x,y) \in S} \| x_t - x \|_2 \leq \gamma$}
	\STATE $\hat{y}_t \gets \tilde{y}$, where $(\tilde{x}, \tilde{y}) = \argmin_{(x,y) \in S} \| x_t - x \|_2$.
\ELSE
	\STATE Predict $\hat{y}_t$ uniformly at random from $\cbr{1,\ldots,K}$, receive feedback $z_t = \one[\hat{y}_t \neq y_t]$.
  \label{line:nn-explore}
	\IF{$z_t = 0$}
    \STATE $S \gets S \cup \cbr{(x_t, \hat{y}_t)}$.
  \ENDIF
\ENDIF
\ENDFOR
\end{algorithmic}
\label{alg:nn}
\end{algorithm}

@@

\begin{lemma}
Suppose $(x_1,y_1), \ldots, (x_T, y_T)$ is a set of $\gamma$-weakly linearly separable examples.
For $i$, $j$ in $\cbr{1,\ldots,T}$, if $x_i$ and $x_j$ have $\ell_2$ distance at most $\gamma$, then $y_i = y_j$.
\label{lem:nn}
\end{lemma}
\begin{proof}
Suppose for the sake on contradiction that $y_i \neq y_j$. By Definition~\ref{definition:weak-linear-separability}, there exists a set of $k$ vectors $(w_1, \ldots, w_k)$, such that Equations~\eqref{equation:weak-linear-separability-1}
and~\eqref{equation:weak-linear-separability-1} are satisfied.

@@

\begin{theorem}
Suppose $(x_1,y_1), \ldots, (x_T, y_T)$ is a set of $\gamma$-weakly linearly separable examples. In addition, for all $t$ in $\cbr{1,\ldots,T}$, $\| x_t \| \leq R$. Then, the expected number of mistakes made by Algorithm~\ref{alg:nn} is such that
\[\Exp[M] \leq (K-1) (\frac R \gamma)^d. \]
\label{thm:margin_at_upper}
\end{theorem}
\begin{proof}
Let $b_t$ be the indicator that line~\ref{line:nn-explore} is executed at timestep $t$,
i.e. we fall into the ``else'' case.
Note that if $b_t = 0$, then by Lemma~\ref{lem:nn}, the prediction $\hat{y}_t$ must equal $y_t$, i.e. $z_t = 0$.
Therefore, $M = \sum_{t=1}^T z_t = \sum_{t=1}^T b_t z_t$. Let $U = \sum_{t=1}^T b_t (1-z_t)$; it can be seen that $U$ is the number of times when $S$ gets updated. As $S$ is a $\gamma$-packing over
$\cbr{x: \|x\|_2 \leq R}$, the size of $S$ is at most $(\frac R \gamma)^d$, which is also an upper bound on $U$.

>>>>

\section{Nearest neighbor algorithm}
\label{section:nearest-neighbor-algorithm}

\begin{algorithm}
\caption{A nearest-neighbor algorithm for \textsc{Online Multiclass Linear Classification with
Bandit Feedback}}
\begin{algorithmic}[1]
\STATE $S \gets \emptyset$.
\FOR{$t=1,2,\ldots,T$:}
\IF{$\min_{(x,y) \in S} \| x_t - x \|_2 \leq \gamma$}
	\STATE $\hat{y}_t \gets \tilde{y}$, where $(\tilde{x}, \tilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}_2$.
\ELSE
	\STATE Predict $\hat{y}_t$ uniformly at random from $\cbr{1,2,\dots,K}$, receive feedback $z_t = \indicator{\hat{y}_t \neq y_t}$.
  \label{line:nn-explore}
	\IF{$z_t = 0$}
    \STATE $S \gets S \cup \cbr{(x_t, \hat{y}_t)}$.
  \ENDIF
\ENDIF
\ENDFOR
\end{algorithmic}
\label{alg:nn}
\end{algorithm}

@@

\begin{lemma}
Suppose $(x_1,y_1), \ldots, (x_T, y_T)$ is a set of $\gamma$-weakly linearly
separable examples. For $i$, $j$ in $\cbr{1,\ldots,T}$, if $x_i$ and $x_j$ have
$\ell_2$ distance at most $\gamma$, then $y_i = y_j$. \label{lem:nn}
\end{lemma}

\begin{proof}
Suppose for the sake on contradiction that $y_i \neq y_j$. By
Definition~\ref{definition:weak-linear-separability}, there exists a set of $k$
vectors $(w_1, \ldots, w_k)$, such that
Equations~\eqref{equation:weak-linear-separability-1}
and~\eqref{equation:weak-linear-separability-1} are satisfied.

@@

\begin{theorem}
\label{thm:margin_at_upper}
Suppose $(x_1,y_1), \ldots, (x_T, y_T)$ is a set of $\gamma$-weakly linearly
separable examples. In addition, for all $t$ in $\cbr{1,2,\dots,T}$, $\| x_t \|
\leq R$. Then, the expected number of mistakes made by Algorithm~\ref{alg:nn} is
such that \[\Exp[M] \le (K-1) (\frac R \gamma)^d. \]
\end{theorem}

\begin{proof}
Let $b_t$ be the indicator that line~\ref{line:nn-explore} is executed at
timestep $t$, i.e. we fall into the ``else'' case. Note that if $b_t = 0$, then
by Lemma~\ref{lem:nn}, the prediction $\hat{y}_t$ must equal $y_t$, i.e. $z_t =
0$. Therefore, $M = \sum_{t=1}^T z_t = \sum_{t=1}^T b_t z_t$. Let $U =
\sum_{t=1}^T b_t (1-z_t)$; it can be seen that $U$ is the number of times when
$S$ gets updated. As $S$ is a $\gamma$-packing over $\cbr{x: \|x\|_2 \leq R}$,
the size of $S$ is at most $(\frac R \gamma)^d$, which is also an upper bound on
$U$.

### 2019-01-22 ICML-submission/appendix-proofs-for-strongly-separable-examples.tex
Fixes

\begin{proof}[Proof of \autoref{theorem:strongly-separable-examples-mistake-upper-bound}]
Let $M = \sum_{t=1}^T z_t$ be the number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes. Let $A = \sum_{t=1}^T \one[S_t \neq \emptyset] z_t$ be the number of
mistakes in the rounds when $S_t \neq \emptyset$, i.e. the number of rounds
line~\ref{line:neg-update} is executed. In addition, let $B = \sum_{t=1}^T
\one[S_t = \emptyset] z_t$ be the number of mistakes in the rounds when $S_t =
\emptyset$. It can be easily seen that $M = A + B$.

Let $C = \sum_{t=1}^T \one[S_t = \emptyset](1 - z_t)$ be the number of rounds
line~\ref{line:pos-update} gets executed. Let $U = \sum_{t=1}^T (\one[S_t \neq
\emptyset] z_t + \one[S_t = \emptyset](1 - z_t))$ be the number of rounds
line~\ref{line:pos-update} or~\ref{line:neg-update} gets executed. In other
words, $U$ is the number of times the $K$-tuple of vectors $(w_1^{(t)},
w_2^{(t)}, \dots, w_K^{(t)})$ gets updated. It can be easily seen that $U = A +
C$.

>>>>

\begin{proof}[Proof of \autoref{theorem:strongly-separable-examples-mistake-upper-bound}]
Let $M = \sum_{t=1}^T z_t$ be the number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes. Let $A = \sum_{t=1}^T \indicator{S_t \neq \emptyset} z_t$ be the number of
mistakes in the rounds when $S_t \neq \emptyset$, i.e. the number of rounds
line~\ref{line:neg-update} is executed. In addition, let $B = \sum_{t=1}^T
\indicator{S_t = \emptyset} z_t$ be the number of mistakes in the rounds when $S_t =
\emptyset$. It can be easily seen that $M = A + B$.

Let $C = \sum_{t=1}^T \indicator{S_t = \emptyset}(1 - z_t)$ be the number of rounds
line~\ref{line:pos-update} gets executed. Let $U = \sum_{t=1}^T (\indicator{S_t \neq
\emptyset} z_t + \indicator{S_t = \emptyset}(1 - z_t))$ be the number of rounds
line~\ref{line:pos-update} or~\ref{line:neg-update} gets executed. In other
words, $U$ is the number of times the $K$-tuple of vectors $(w_1^{(t)},
w_2^{(t)}, \dots, w_K^{(t)})$ gets updated. It can be easily seen that $U = A +
C$.
